---
title: "Energy_Consumption"
author: "Luis Varela"
date: "7 01 2020"
output: html_document
---

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Introduction</h2>

<p style="text-align: justify;">There are a series of Submeters that meassure the energy consumption in different parts of an appartment. These meassurements will be used to forecast consumption on the next year, but above all there will be an exploration that will be made to give recommendations to the tenant of the building on how to save energy and why these devices can help reduce the energy consumption.</p>

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Used Libraries</h2>


```{r, warning=FALSE, message=FALSE}
library(RMySQL)     #for the SQL queries
library(DBI)        #necessary for the RMySQL library
library(ggplot2)    #graph library
library(plotly)     #this one is better for the line graphs
library(knitr)      #for the tables aes
library(magrittr)   #magic %>% 
library(dplyr)      #data wrangling
library(lubridate)  #convert date and time
library(grid)       #formating the ggplot graphics
library(ggfortify)  #ploting TS object
library(forecast)   #for forecasting
```


```{r, include=FALSE}
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold.", type)]])) return(res)
    
    paste0(
      "<details><summary>", type, "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot")
)
```

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Fetch Data and Select Variables</h2>

```{r, warning=FALSE, message=FALSE}
#Importing the data, create a database connection

con <- dbConnect(MySQL(),
                 user = 'deepAnalytics',
                 password = 'Sqltask1234!',
                 dbname = 'dataanalytics2018',
                 host = 'data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')

#List the tables contained in the database 

dbListTables(con)

#quick check of attributes listed in one of the lists, I kow beforehand they all, except one, contain the same attributes

dbListFields(con, 'yr_2006')

#save all df locally with all the info in them

yr2006_local <- dbGetQuery(con, "SELECT * FROM yr_2006")
write.csv(yr2006_local, file = "Data/yr2006_local")

yr2007_local <- dbGetQuery(con, "SELECT * FROM yr_2007")
write.csv(yr2007_local, file = "Data/yr2007_local")

yr2008_local <- dbGetQuery(con, "SELECT * FROM yr_2008")
write.csv(yr2008_local, file = "Data/yr2008_local")

yr2009_local <- dbGetQuery(con, "SELECT * FROM yr_2009")
write.csv(yr2009_local, file = "Data/yr2009_local")

yr2010_local <- dbGetQuery(con, "SELECT * FROM yr_2010")
write.csv(yr2010_local, file = "Data/yr2010_local")

#for the analysis I need to implement, only 5 attributes are relevant, Time, Date, Global_active_power and the 3 Sub_metering

yr_2006 <- select(yr2006_local, Date, Time, Sub_metering_1, Sub_metering_2, Sub_metering_3, Global_active_power)
  
yr_2007 <- select(yr2007_local, Date, Time, Sub_metering_1, Sub_metering_2, Sub_metering_3, Global_active_power)

yr_2008 <- select(yr2008_local, Date, Time, Sub_metering_1, Sub_metering_2, Sub_metering_3, Global_active_power)

yr_2009 <- select(yr2009_local, Date, Time, Sub_metering_1, Sub_metering_2, Sub_metering_3, Global_active_power)
  
yr_2010 <- select(yr2010_local, Date, Time, Sub_metering_1, Sub_metering_2, Sub_metering_3, Global_active_power)
```

```{r, echo=FALSE, results='hide'}
#disconnect from server

dbDisconnect(con)
```

```{r, collapse=TRUE}
#str() visualized as kable table

x_str <- function(data){
 data.frame(
            variable = names(data),
            class = sapply(data, typeof),
            first_values = sapply(data, function(x) paste0(head(x), collapse = ", ")),
            row.names = NULL) %>%
            kable()
}
```

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Exploration of the Data</h2>

## {.tabset .tabset-fade}

### Year 2006
```{r, fold.plot=FALSE}
x_str(yr_2006)
kable(summary(yr_2006))
kable(head(yr_2006))
```

### Year 2007
```{r, fold.plot=FALSE}
x_str(yr_2007)
kable(summary(yr_2007))
kable(head(yr_2007))
```

### Year 2008
```{r, fold.plot=FALSE}
x_str(yr_2008)
kable(summary(yr_2008))
kable(head(yr_2008))
```

### Year 2009
```{r, fold.plot=FALSE}
x_str(yr_2009)
kable(summary(yr_2009))
kable(head(yr_2009))
```

### Year 2010
```{r, fold.plot=FALSE}
x_str(yr_2010)
kable(summary(yr_2010))
kable(head(yr_2010))
```

## {-}

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Data Manipulation</h2>


```{r, fold.plot=FALSE}
#join the df that are complete and will be used for the analysis

yr_07_08_09 <- bind_rows(yr_2007, yr_2008, yr_2009)

#change the names of the columns

names(yr_07_08_09)[names(yr_07_08_09) == "Sub_metering_1"] <- "Kitchen"
names(yr_07_08_09)[names(yr_07_08_09) == "Sub_metering_2"] <- "Laundry_room"
names(yr_07_08_09)[names(yr_07_08_09) == "Sub_metering_3"] <- "Water_heater"

#changing the submeters to kw/h

yr_07_08_09$Kitchen <- yr_07_08_09$Kitchen/1000
yr_07_08_09$Laundry_room <- yr_07_08_09$Laundry_room/1000
yr_07_08_09$Water_heater <- yr_07_08_09$Water_heater/1000
kable(head(yr_07_08_09, n = 5))

#change the reading to the same unit in the "Global_active_power" to kw/h, and take from the reading the sum of the other three sub meters

yr_07_08_09$Rest_house <- round(yr_07_08_09$Global_active_power/60 - (yr_07_08_09$Kitchen + yr_07_08_09$Laundry_room + yr_07_08_09$Water_heater), 3)
kable(head(yr_07_08_09))

#combine date and time attributes in a new column

yr_07_08_09$Date_time <- paste(yr_07_08_09$Date, yr_07_08_09$Time)

#moving the column

yr_07_08_09 <- yr_07_08_09[,c(ncol(yr_07_08_09), 1:(ncol(yr_07_08_09)-1))]
```

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Creation of New Variables Based on Date_time</h2>

```{r}
#convert Date_time to POSIXct

yr_07_08_09$Date_time <- ymd_hms(yr_07_08_09$Date_time)
yr_07_08_09$Time <- hms(yr_07_08_09$Time)
yr_07_08_09$Date <- ymd(yr_07_08_09$Date)

#add the time zone

attr(yr_07_08_09$Date_time, "tzone") <- "Europe/Paris"

## Inspect the data types

x_str(yr_07_08_09)

#create new columns for information mining

yr_07_08_09$Year      <- year(yr_07_08_09$Date_time)
yr_07_08_09$Quarter   <- quarter(yr_07_08_09$Date_time)
yr_07_08_09$Weekday   <- wday(yr_07_08_09$Date_time)
yr_07_08_09$Minute    <- minute(yr_07_08_09$Date_time)
yr_07_08_09$Month     <- month(yr_07_08_09$Date_time)
yr_07_08_09$Day       <- day(yr_07_08_09$Date_time)
yr_07_08_09$Hour      <- hour(yr_07_08_09$Date_time)
```


```{r, include=FALSE}
#theme for the ggplot below

blank_theme <- theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_text(size = 14, face = "bold")
  )
```


```{r, out.width=c('50%', '50%'), fig.show='hold', warning=FALSE, message=FALSE, fold.plot=FALSE}
#checking how much energy each submeter used in 2008 during the weekend and workdays

submeters_consumption_08 <- yr_07_08_09 %>%
                             filter(Year == 2008, Weekday == 1 & 2 & 3 & 4 & 5) %>%
                             select(Kitchen, Laundry_room, Water_heater, Rest_house) %>% 
                             summarise_all(funs(sum)) %>% 
                             tidyr::gather(key = Area, value = KwH) %>% 
                             mutate(Porcentage = round(KwH/sum(KwH) * 100, 2))

#pie chart showing the porcentage of power used by each submeter in 2008

ggplot(submeters_consumption_08, aes(x = "", y = KwH, fill = Area)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Dark2") +
  blank_theme +   #Previously created object, refer to RMD doc to see it
  theme(axis.text.x = element_blank()) +
#  theme(plot.margin = unit(c(0, -10, -10, -30), "mm")) +
  geom_text(aes(label = Porcentage), position = position_stack(vjust = 0.5)) +
  ggtitle("Sum of the Energy Consumption During all Weekdays in 2008")

#checking how much energy each submeter used during the sum of the weekends of the year

submeters_consumption_weekends <- yr_07_08_09 %>%
                                   filter(Weekday == 6 & 7) %>%
                                   select(Kitchen, Laundry_room, Water_heater, Rest_house) %>% 
                                   summarise_all(funs(sum)) %>% 
                                   tidyr::gather(key = Area, value = KwH) %>% 
                                   mutate(Porcentage = round(KwH/sum(KwH) * 100, 2))

#pie chart showing the porcentage of power used by each submeter in 2008

ggplot(submeters_consumption_weekends, aes(x = "", y = KwH, fill = Area)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Dark2") +
  blank_theme +   #Previously created object, refer to RMD doc to see it
  theme(axis.text.x = element_blank()) +
#  theme(plot.margin = unit(c(0, -10, -10, -30), "mm")) +
  geom_text(aes(label = Porcentage), position = position_stack(vjust = 0.5)) +
  ggtitle("Sum of the Energy Consumption During All Weekends in 2008")

#Note to self: quite the jump in hot water ussage during the weekends. People are more at home, that´d explain it.
#overall ussage of energy is 6 times higher on the weekends than on the weekdays
```

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Inspection Patterns of Consumption in Time</h2>


## {.tabset .tabset-fade}

### DAY
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.fullwidth = TRUE, fig.width=9.5}
#plot of the mid summer week, friday 2009

Fri_summer <- yr_07_08_09 %>% 
                      filter(Year == 2009 & Month == 6 & Day == 26 & (Minute == 0 | Minute == 10 | Minute == 20 | Minute == 30 | Minute == 40 | Minute == 50))

plot_ly(Fri_summer, x = ~Date_time, y = ~Kitchen, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
  add_trace(y = ~Laundry_room, mode = 'lines') %>%
  add_trace(y = ~Water_heater, mode = 'lines') %>%
  add_trace(y = ~Rest_house, mode = 'lines') %>%
  layout(title = "Energy Consumption Mid-Summer Friday 26th of June 2009",
         xaxis = list(title = "Time"),
         yaxis = list (title = "KwH"))
```

### WEEK
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.fullwidth = TRUE, fig.width=9.5}
#plot of the mid summer week of 2009

Fri_summer <- yr_07_08_09 %>% 
                  filter(Date_time > "2009-06-22" & Date_time < "2009-06-28" & (Hour == 1 | Hour == 2 | Hour == 3 | Hour == 4 | Hour == 5 | Hour                          == 6 | Hour == 7 | Hour == 8 | Hour == 9 | Hour == 10 | Hour == 11 | Hour == 12 | Hour == 13 | Hour == 14 | Hour == 15                          | Hour == 16 | Hour == 17 | Hour == 18 | Hour == 19 | Hour == 20 | Hour == 21 |Hour == 22 | Hour == 23 | Hour == 0))

plot_ly(Fri_summer, x = ~Date_time, y = ~Kitchen, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
  add_trace(y = ~Laundry_room, mode = 'lines') %>%
  add_trace(y = ~Water_heater, mode = 'lines') %>%
  add_trace(y = ~Rest_house, mode = 'lines') %>%
  layout(title = "Energy Consumption Mid-Summer Week 2009",
         xaxis = list(title = "Time"),
         yaxis = list (title = "KwH"))
```

### MONTH
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.fullwidth = TRUE, fig.width=9.5}
#plot of the mid summer week of 2009

Jul_2009 <- yr_07_08_09 %>% 
                  filter(Year == 2009 & Month == 7 & (Hour == 1 | Hour == 2 | Hour == 3 | Hour == 4 | Hour == 5 | Hour == 6 | Hour == 7 | Hour                            == 8 | Hour == 9 | Hour == 10 | Hour == 11 | Hour == 12 | Hour == 13 | Hour == 14 | Hour == 15 | Hour == 16 | Hour ==                           17 | Hour == 18 | Hour == 19 | Hour == 20 | Hour == 21 |Hour == 22 | Hour == 23 | Hour == 0))

plot_ly(Jul_2009, x = ~Date_time, y = ~Kitchen, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
  add_trace(y = ~Laundry_room, mode = 'lines') %>%
  add_trace(y = ~Water_heater, mode = 'lines') %>%
  add_trace(y = ~Rest_house, mode = 'lines') %>%
  layout(title = "Energy Consumption June 2009",
         xaxis = list(title = "Time"),
         yaxis = list (title = "KwH"))
```

## {-}

<div style="margin-bottom:100px;">
</div>

<h2 style="text-align: center;">Forecasting on Each Submeter</h2>


## {.tabset .tabset-fade}

### KITCHEN
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
#subset to one observation per day on 

weekly_07_09 <- filter(yr_07_08_09, Weekday == 1 & 
                              Hour == 20 & Minute == 1)

#TS object with submeter from Kitchen

tsKitchen <- ts(weekly_07_09$Kitchen, frequency=52, start = c(2007, 1))

#ploting submeter Kitchen with autoplot - add labels, color

autoplot(tsKitchen, ts.colour = 'red', xlab = "Time", ylab = "KwH", main = "Kitchen")

#trying other plotting package

plot.ts(tsKitchen)

#linear model
fitKitchen <- tslm(tsKitchen ~ trend + season) 
summary(fitKitchen)

#create the forecast for submeter in the Kitchen. Forecast ahead 20 time periods 

forecastfitKitchen <- forecast(fitKitchen, h=20)
plot(forecastfitKitchen)

#create submeter Kitchen forecast with confidence levels 80 and 90

forecastfitKitchen2 <- forecast(fitKitchen, h=20, level = c(80, 90))
plot(forecastfitKitchen2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time")

#ecompose submeter Kitchen into trend, seasonal and remainder

decomposedForecast <- decompose(tsKitchen)

#plot decomposed sub-meter 1 

plot(decomposedForecast, col = "red")

#statistic summary

summary(decomposedForecast)

#seasonal adjusting sub-meter 1 by subtracting the seasonal component & plot

tsAdjusted <- tsKitchen - decomposedForecast$seasonal
autoplot(tsAdjusted, col = 'red', xlab = "Time", ylab = "KwH", main = "Kitchen")

#now usging Holt Winters Exponential Smoothing

tsHW <- HoltWinters(tsAdjusted, beta = FALSE, gamma = FALSE)
plot(tsHW, ylim = c(-0.035, 0.045))

#forecast Holt Winters

tsHWforecast <- forecast(tsHW, h=25)
plot(tsHWforecast, ylim = c(0, 0.04), ylab= "KwH", xlab = "Time - Kitchen")

#forecast HoltWinters with diminished confidence levels

tsHWforecast2 <- forecast(tsHW, h = 25, level = c(10, 25))
plot(tsHWforecast2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time - Kitchen", start(2010))
```


### LAUNDRY-ROOM
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
#subset to one observation per day on 

weekly_07_09_laundry <- filter(yr_07_08_09, Weekday == 1 & 
                              Hour == 20 & Minute == 1)

#TS object with submeter from Kitchen

tsLaundry <- ts(weekly_07_09_laundry$Laundry_room, frequency=52, start = c(2007, 1))

#ploting submeter Kitchen with autoplot - add labels, color

autoplot(tsLaundry, ts.colour = 'red', xlab = "Time", ylab = "KwH", main = "Laundry_room")

#trying other plotting package

plot.ts(tsLaundry)

#linear model
fitLaundry <- tslm(tsLaundry ~ trend + season) 
summary(fitLaundry)

#create the forecast for submeter in the Kitchen. Forecast ahead 20 time periods 

forecastfitLaundry <- forecast(fitLaundry, h=20)
plot(forecastfitKitchen)

#create submeter Kitchen forecast with confidence levels 80 and 90

forecastfitLaundry2 <- forecast(fitLaundry, h=20, level = c(80, 90))
plot(forecastfitLaundry2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time")

#ecompose submeter Kitchen into trend, seasonal and remainder

decomposedForecastLaundry <- decompose(tsLaundry)

#plot decomposed sub-meter 1 

plot(decomposedForecastLaundry, col = "red")

#statistic summary

summary(decomposedForecastLaundry)

#seasonal adjusting sub-meter 1 by subtracting the seasonal component & plot

tsAdjustedLaundry <- tsLaundry - decomposedForecastLaundry$seasonal
autoplot(tsAdjustedLaundry, col = 'red', xlab = "Time", ylab = "KwH", main = "Laundry_Room")

#now usging Holt Winters Exponential Smoothing

tsHWLaundry <- HoltWinters(tsAdjustedLaundry, beta = FALSE, gamma = FALSE)
plot(tsHW, ylim = c(-0.035, 0.045))

#forecast Holt Winters

tsHWforecastLaundry <- forecast(tsHWLaundry, h=25)
plot(tsHWforecast, ylim = c(0, 0.04), ylab= "KwH", xlab = "Time - Kitchen")

#forecast HoltWinters with diminished confidence levels

tsHWforecastLaundry2 <- forecast(tsHWLaundry, h = 25, level = c(10, 25))
plot(tsHWforecastLaundry2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time - Laundry_Room", start(2010))
```


### WATER HEATER
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
#subset to one observation per day on 

weekly_07_09_WaterH <- filter(yr_07_08_09, Weekday == 1 & 
                              Hour == 20 & Minute == 1)

#TS object with submeter from Kitchen

tsWaterH <- ts(weekly_07_09_WaterH$Water_heater, frequency=52, start = c(2007, 1))

#ploting submeter Kitchen with autoplot - add labels, color

autoplot(tsWaterH, ts.colour = 'red', xlab = "Time", ylab = "KwH", main = "Water_heater and AC")

#trying other plotting package

plot.ts(tsWaterH)

#linear model
fitWaterH <- tslm(tsWaterH ~ trend + season) 
summary(fitWaterH)

#create the forecast for submeter in the Kitchen. Forecast ahead 20 time periods 

forecastfitWaterH <- forecast(fitWaterH, h=20)
plot(forecastfitWaterH)

#create submeter Kitchen forecast with confidence levels 80 and 90

forecastfitWaterH2 <- forecast(fitWaterH, h=20, level = c(80, 90))
plot(forecastfitWaterH2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time")

#ecompose submeter Kitchen into trend, seasonal and remainder

decomposedForecastWaterH <- decompose(tsWaterH)

#plot decomposed sub-meter 1 

plot(decomposedForecastWaterH, col = "red")

#statistic summary

summary(decomposedForecastWaterH)

#seasonal adjusting sub-meter 1 by subtracting the seasonal component & plot

tsAdjustedWaterH <- tsWaterH - decomposedForecastWaterH$seasonal
autoplot(tsAdjustedWaterH, col = 'red', xlab = "Time", ylab = "KwH", main = "Water_heater and AC")

#now usging Holt Winters Exponential Smoothing

tsHWWaterH <- HoltWinters(tsAdjustedWaterH, beta = FALSE, gamma = FALSE)
plot(tsWaterH, ylim = c(-0.035, 0.045))

#forecast Holt Winters

tsHWforecastWaterH <- forecast(tsHWWaterH, h=25)
plot(tsHWforecastWaterH, ylim = c(0, 0.04), ylab= "KwH", xlab = "Time - Water_heater and AC")

#forecast HoltWinters with diminished confidence levels

tsHWforecastWaterH2 <- forecast(tsHWWaterH, h = 25, level = c(10, 25))
plot(tsHWforecastWaterH2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time - Water_heater and AC", start(2010))
```


### REST OF THE HOUSE
```{r, fold.plot=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
#subset to one observation per day on 

weekly_07_09_House <- filter(yr_07_08_09, Weekday == 1 & 
                              Hour == 20 & Minute == 1)

#TS object with submeter from Kitchen

tsHouse <- ts(weekly_07_09_House$Rest_house, frequency=52, start = c(2007, 1))

#ploting submeter Kitchen with autoplot - add labels, color

autoplot(tsHouse, ts.colour = 'red', xlab = "Time", ylab = "KwH", main = "Rest of the House")

#trying other plotting package

plot.ts(tsHouse)

#linear model
fitHouse <- tslm(tsHouse ~ trend + season) 
summary(fitHouse)

#create the forecast for submeter in the Kitchen. Forecast ahead 20 time periods 

forecastfitHouse <- forecast(fitHouse, h=20)
plot(forecastfitHouse)

#create submeter Kitchen forecast with confidence levels 80 and 90

forecastfitHouse2 <- forecast(fitHouse, h=20, level = c(80, 90))
plot(forecastfitHouse2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time")

#ecompose submeter Kitchen into trend, seasonal and remainder

decomposedForecastHouse <- decompose(tsHouse)

#plot decomposed sub-meter 1 

plot(decomposedForecastHouse, col = "red")

#statistic summary

summary(decomposedForecastHouse)

#seasonal adjusting sub-meter 1 by subtracting the seasonal component & plot

tsAdjustedHouse <- tsHouse - decomposedForecastHouse$seasonal
autoplot(tsAdjustedHouse, col = 'red', xlab = "Time", ylab = "KwH", main = "Rest of the House")

#now usging Holt Winters Exponential Smoothing

tsHWHouse <- HoltWinters(tsAdjustedHouse, beta = FALSE, gamma = FALSE)
plot(tsHouse, ylim = c(-0.035, 0.045))

#forecast Holt Winters

tsHWforecastHouse <- forecast(tsHWHouse, h=25)
plot(tsHWforecastHouse, ylim = c(0, 0.04), ylab= "KwH", xlab = "Time - Rest of the House")

#forecast HoltWinters with diminished confidence levels

tsHWforecastHouse2 <- forecast(tsHWHouse, h = 25, level = c(10, 25))
plot(tsHWforecastHouse2, ylim = c(0, 0.04), ylab = "KwH", xlab = "Time - Rest of the House", start(2010))
```


## {-}

```{r}
#sesh info
sessionInfo()
```